---
title: "Problem Set 3"
author: "MaCCS 201 - Fall 2025"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
rm(list = ls()) # clear memory


```

### Please submit markdown file named [last_name]_[first_name]_ps1.Rmd or a pdf with all code and answers. 

##Part 1: Blackburn and Neumark (1992)

Let's take a look at the data from Blackburn and Neumark (QJE 1992). 
 We would like to estimate the model: 
 
\begin{eqnarray}
\log(wage) &=& \beta_o + exper\cdot\beta_1 +tenure\cdot\beta_2+
married\cdot\beta_3+ south\cdot\beta_4 + \\ && urban\cdot\beta_5+
black\cdot\beta_6+ educ\cdot\beta_7 + abil\cdot\gamma +
\varepsilon \nonumber
\end{eqnarray}

One of the big problems in the wage literature is that
we do not observe ability. If ability is not
correlated with any of the right hand side variables, we can
include it in the disturbance and nothing is lost by not observing
it. If, however, it is correlated with one or more of the right
hand side variables, OLS is no longer unbiased or consistent.
Assume that ability is correlated with education and none of the
other right hand side variables. 

a) Derive the bias of $\beta_7$ and
show what direction the bias goes in depending on whether the
correlation between ability and education is positive or negative.


b)  You showed in the first part that we can derive the sign/direction
of the bias. One approach that has been take in the literature is
using a "proxy" variable for the unobservable ability. We will use
IQ here to proxy for ability. Estimate the model above excluding
ability, record your parameter estimates, standard errors and
$R^2$. 

```{r q1b, include=TRUE}


```


c) Estimate the model including IQ as a proxy, record
your parameter estimates, standard errors and $R^2$.

```{r q1c, include=TRUE}

```

d)  What happens to returns to schooling? Does this result
confirm your suspicion of how ability and schooling are expected
to be correlated?  



## Part 2: Card (1995)

a) Read the data into R. Plot the series make sure your data are
read in correctly. 

```{r q1d, include=TRUE}

```
b)  Estimate a $\log(wage)$ regression via
Least Squares with $educ$, $exper$, $exper^2$, $black$, $south$,
$smsa$, $reg661$ through $reg668$ and $smsa66$ on the right hand
side. Check your results against Table2, column 5.

```{r q1e, include=TRUE}

```


c) Estimate a
reduced form equation for $educ$ containing all of the explanatory
variables and the dummy variable $nearc4$. Is the partial
correlation between $nearc4$ and $educ$ statistically significant?

```{r q1f, include=TRUE}

```

d)  Estimate the $\log(wage)$ equation by instrumental
variables, using $nearc4$ as an instrument for $educ$. Compare the
95\% confidence interval for the return to education to that
obtained from the Least Squares regression above. 

```{r q1g, include=TRUE}

```


e) (Bonus) Now use
multiple instruments. Use $nearc2$ and $nearc4$ as instruments for
$educ$. Comment on the significance of the partial correlations of
both instruments in the reduced form. 

```{r q1h, include=TRUE}

```


 



